---
title: "p8105_hw6_tj2519"
output: github_document
date: "2023-11-29"
---

```{r}
library(tidyverse)
library(p8105.datasets)
set.seed(1)
```

# Problem 2
## Initial setup (load the dataset) and create a bootstrap function
```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2022-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())

boot_sample = function(df) {
  sample_frac(df, replace = TRUE)
}

bootboot_straps = 
  tibble(strap_number = 1:10) |> 
  mutate(
    strap_sample = map(strap_number, \(df) boot_sample(weather_df))
  )

```


## Run Linear Regression to the samples and calculate the log(b1*b2) and r^2 for each bootstrap sample
```{r}
bootboot_straps_results =
  bootboot_straps %>% 
  mutate(
    models = map(strap_sample, \(df) lm(tmax ~ tmin + prcp, data = df)),
    results = map(models, broom::tidy),
    r_sq = map(models, broom::glance),
    log_value = map_dbl(results, ~log(.x[[2]][[2]] * .x[[2]][[3]]))) %>% 
  mutate(r_sq = map(r_sq, ~ .x %>% select(r.squared))) %>% 
  select(-strap_sample, -models) %>% 
  unnest(results, r_sq) %>% 
  janitor::clean_names()
  
```

## Plot the estimate distributions (r^2 and log(beta1*beta2))
```{r}
bootboot_straps_results %>% 
  ggplot(aes(x = r_squared)) + 
  geom_density() +
  labs(title = "Distribution of r_squared Across the Samples")


bootboot_straps_results %>% 
  ggplot(aes(x = log_value)) + 
  geom_density() +
  labs(title = "Distribution of log(beta1*beta2) Across the Samples")
  
```

## Compute 95% CI for r^2 and log(beta1*beta2)
```{r}
bootboot_straps_results %>% 
  summarize(r_sq_ci_lower = quantile(r_squared, 0.025),
         r_sq_ci_upper = quantile(r_squared, 0.975),
         log_ci_lower = quantile(log_value, 0.025, na.rm = TRUE),
         log_ci_upper = quantile(log_value, 0.975, na.rm = TRUE)) %>% 
  knitr::kable()
```



# Problem 3
## Load the dataset and clean it, drop any NA values, so it would be easier in later building regression model parts
```{r}
problem3_df = 
  read_csv("data/birthweight.csv") %>% 
  janitor::clean_names() %>% 
  drop_na() %>% 
  mutate(babysex = recode(babysex, `1` = "male",
                          `2` = "female"),
         frace = recode(frace, `1` = "White", 
                        `2` = "Black",
                        `3` = "Asian",
                        `4` = "Puerto Rican",
                        `8` = "Other",
                        `9` = "Unknown"),
         mrace = recode(frace, `1` = "White", 
                        `2` = "Black",
                        `3` = "Asian",
                        `4` = "Puerto Rican",
                        `8` = "Other"),
         malform = recode(malform, `0` = "absent",
                          `1` = "present"),
         babysex = fct_relevel(babysex, c("male", "female")),
         frace = fct_infreq(frace),
         mrace = fct_infreq(mrace),
         malform = fct_relevel(malform, c("absent", "present")),
         ) 
  

```

## Create own model using stepwise method. 
### In this method, I used 'step' function in R. By using the stepwise method, the algorithm would build a linear regression model based on all the variables available in the dataset. By default, step function build the final model based on both the forward (start with intercept only) and backward method (start with all the covariates), till the default criterion - AIC value of the model could not be improved.
```{r}
crude_model = lm(bwt ~ ., data = problem3_df)
final_model = step(initial_model)

model_results = summary(final_model) %>% 
    broom::tidy()
```



## Run Linear regression with length at birth and gestational age as predictors (main effects only)
```{r}
model_1 = lm(bwt ~ blength + gaweeks, data = problem3_df) %>% 
  broom::tidy()
  
model_2 = lm(bwt ~ bhead * blength * babysex, data = problem3_df) %>% 
  broom::tidy()

```


